\clearpage


\chapter{Summary and Outlook}

The investigation of glass transition temperature (\(T_g\)) in thin polymer films has been a subject of considerable debate, particularly when comparing the behavior of these films to their bulk counterparts. This research delves into this debate by examining the complex dynamics of stiff polymers, specifically Poly-TPD and Poly-Bisphenol, under conditions of confinement. Through a focused analysis on how interface-induced changes affect these polymers, the present work contributes to a deeper understanding of the mechanisms governing the glass transition and relaxation processes in thin films.

Many studies on thin polymer films have predominantly focused on flexible polymers, which have short segmental lengths, like polystyrene \cite{INOUE2006367}. In these polymers, the effects induced by the interface typically dissipate rapidly as one moves away from the boundary, resulting in a relatively shallow penetration depth of these effects \cite{winkler2021}, and thus an indifference in the \(T_g\) and other properties as the thickness of the film changes (although that is still subject of debate). However, the present work adopts a different approach by focusing on the stiff polymers, namely Poly-TPD and P-Bisphenol. The inherent stiffness of these polymers, characterized by longer segmental lengths, enables a deeper penetration of the interface effects into the film, thus "artificially" enhancing the resolution of these effects.

By using polymers with greater chain stiffness, this research demonstrates that the effects of confinement and interfacial interactions can penetrate more deeply into the film, providing a more detailed and nuanced understanding of the dynamics at play. The findings indicate that the \(T_g\) and relaxation processes in these stiff polymer films differ markedly from those observed in thin films of shorter-segment polymers, as well as from bulk materials, underscoring the critical role of polymer chain stiffness in determining the glass transition behavior in confined systems.

What can be drawn from this analysis is that the correspondence between bulk vs. confined polymer properties is not always applicable, especially when it comes to longer, stiffer-segment polymers like those studied here. Special attention is required, as interfacial effects can be quite prominent, as observed in Poly-TPD. This phenomenon is also expected for P-Bisphenol, though the data at hand was slightly more challenging to interpret, and further experiments may be needed for full confirmation.

The implications of this research also extend beyond the specific polymers studied, highlighting the need to consider the unique dynamics of stiff polymers in confined geometries. The deeper penetration of interface-induced effects in stiff polymers like Poly-TPD and Poly-Bisphenol offers a different perspective on the glass transition and relaxation phenomena in thin films. This understanding is crucial for optimizing the performance of polymer-based thin film technologies, particularly in applications where precise control over thermal and mechanical properties is essential.

Furthermore, with novel analytical techniques, such as those explored in recent works, it is now possible to deconvolute the depth gradient of the relaxation times across the polymer films. The study by Cheng et al. \cite{cheng2024} has shown how these gradients can be characterized using advanced dielectric spectroscopy methods. The findings from this research provide a valuable data set for such deconvolution efforts, offering a promising path for future work to resolve the spatial gradient of relaxation dynamics with greater precision, thereby deepening our understanding of how interface effects manifest across different depths in confined polymer systems.

% Example citation:
% \cite{cheng2024}

%% Lade offizielle Zusammenfassung
\input{official/zusammenfassung}

\noindent
%% Beginn der eigenen Zusammenfassung
